---
title: "Logistic_regression_Titanic"
author: "Marta Ortigas"
date: "16/11/2022"
output: html_document
---
First, we will see if our data fits a logistic regression model. Then, we will
use logistic regression to find out which explanatory variables are more significant
in the survival on the Titanic.
For this part of the project we will need the packages pscl and caret.

To get started, we observe the data:
```{r}
head(titanic)
```
As we can see, our response variable, survive the Titanic (X2urvived), is binary
which is perfect when doing a logistic model.

We do a general logistic regression to see which explanatory variables explain
better our response variable (survival on the Titanic).
```{r}
general_logisitc <- glm(X2urvived ~ ., data = titanic, family = "binomial")
summary(general_logisitc)
```
P-values:
Age: 5.35e-07
Sex: < 2e-16
Pclass: 1.06e-15

As we can see, the explanatory variables with a better significance level are Age,
Sex and Pclass since they have low p-values (we do not count Passengerid because
is not relevant).
Also, we get that:
log(pi/(1-pi)) = 3.2061733 - 0.0343026 * age + 2.2640321 * sex - 0.9859739 * Pclass

To continue, we will see the importance of each predictor in the model using the
function varImp() from the package caret.
```{r}
caret::varImp(general_logisitc)
```
Higher values indicate more importance, so the results obtained match up very well
with the p-values from our model. Being Age, Sex and Pclass good predictors.

Next, we are going to calculate McFadden’s R² to assess how well our model
fits the data.
The value ranges from 0 to just under 1, with higher values indicating better model fit.
```{r}
pscl::pR2(general_logisitc)["McFadden"]
```
Our McFadden’s R² is equal 0.3316875 which means that the logistic model fits
pretty well our data.


Secondly, we will regress our response variable in this explanatory ones and then we will
use the logit function to determine pi = proportion of 1's (survive/success) at any X.

SURVIVAL & AGE
```{r}
logisitc_1 <- glm(X2urvived ~ Age, data = titanic, family = "binomial")
summary(logisitc_1)
```
We get that: log(pi/(1-pi)) = -0.746635 - 0.010059 * age 
How the age can vary very widely we are not going to compute the probability of
surviving in each one of the possible ages. 

Now let's plot the fitted logistic model:
```{r}
plot(titanic$Age, titanic$X2urvived, xlab = "Age", ylab = "survivance", ylim=c(-0.1,1.1))
curve(predict(logisitc_1,data.frame(Age=x),type = "resp"),add=TRUE,col="blue",lwd=2)
points(titanic$Age,fitted(logisitc_1),pch=1,col="blue")
```
In blue we have represented the predicted values according to the model logistic_1.



SURVIVAL & GENDER
```{r}
logisitc_2 <- glm(X2urvived ~ Sex, data = titanic, family = "binomial")
summary(logisitc_2)
```
We get that: log(pi/(1-pi)) = -1.9072 + 1.9072 * sex 
Remember that women = 1, men = 0, so we have that:

For women: 
log(pi/(1-pi)) = -1.9072 + 1.9072 * 1 = 0 → exp(log(pi/(1-pi))) = exp(0) →
→ pi/(1-pi) = exp(0) → pi/(1-pi) = 1 → pi = 1(1-pi) → pi = 1-pi >
→ 2pi = 1 → pi = 1/2 → pi = 0.5

The probability of surviving the Titanic being a woman was 50% .


For men:
log(pi/(1-pi)) = -1.9072 + 1.9072 * 0 = -1.9072 → exp(log(pi/(1-pi))) = exp(-1.9072) →
→ pi/(1-pi) = exp(-1.9072) → pi/(1-pi) = 0.1484956 → pi = 0.1484956(1-pi) →
pi = 0.1484956-0.1484956pi → 1.148496pi = 0.1484956 → pi = 0.1484956/1.148496 →
→ pi = 0.1292957

The probability of surviving the Titanic being a man was 12.93%. Let's say 13%.

Now let's plot the fitted logistic model:
```{r}
plot(titanic$Sex, titanic$X2urvived, xlab = "Sex", ylab = "survivance", ylim=c(-0.1,1.1))
curve(predict(logisitc_2,data.frame(Sex=x),type = "resp"),add=TRUE,col="blue",lwd=2)
points(titanic$Sex,fitted(logisitc_2),pch=1,col="blue")
```
In blue we have represented the predicted values according to the model logistic_2.



SURVIVAL & CLASS
```{r}
logisitc_3 <- glm(X2urvived ~ Pclass, data = titanic, family = "binomial")
summary(logisitc_3)
```
We get that: log(pi/(1-pi)) = 0.37750 - 0.64652 * Pclass 
Remember that Upper = 1, Middle = 2 and Lower = 3, so we have that:

For Lower class:
log(pi/(1-pi)) = 0.37750 - 0.64652 * 3 = -1.56206 → exp(log(pi/(1-pi))) = exp(-1.56206) →
→ pi/(1-pi) = exp(-1.56206) → pi/(1-pi) = 0.2097036 → pi = 0.2097036(1-pi) → 
→ pi = 0.2097036 - 0.2097036pi → 1.209704pi = 0.2097036 → pi = 0.2097036/1.209704 → 
→ pi = 0.1733512

The probability of surviving the Titanic traveling in lower class was 17.34%.


For Middle class:
log(pi/(1-pi)) = 0.37750 - 0.64652 * 2 = -0.91554 → exp(log(pi/(1-pi))) = exp(-0.91554) →
→ pi/(1-pi) = exp(-0.91554) → pi/(1-pi) = 0.4003004 → pi = 0.4003004(1-pi) → 
→ pi = 0.4003004 - 0.4003004pi → 1.4003pi = 0.4003004 → pi = 0.4003004/1.4003 → 
→ pi = 0.2858676

The probability of surviving the Titanic traveling in middle class was 28.59%.


For Upper class:
log(pi/(1-pi)) = 0.37750 - 0.64652 * 1 = -0.26902 → exp(log(pi/(1-pi))) = exp(-0.26902) →
→ pi/(1-pi) = exp(-0.26902) → pi/(1-pi) = 0.764128 → pi = 0.764128(1-pi) → 
→ pi = 0.764128 - 0.764128pi → 1.764128pi = 0.764128 → pi = 0.764128/1.764128 → 
→ pi = 0.4331477

The probability of surviving the Titanic traveling in upper class was 43.31%.

Let's plot the fitted logistic model:
```{r}
plot(titanic$Pclass, titanic$X2urvived, xlab = "Pclass", ylab = "survivance", ylim=c(-0.1,1.1))
curve(predict(logisitc_3,data.frame(Pclass=x),type = "resp"),add=TRUE,col="blue",lwd=2)
points(titanic$Pclass,fitted(logisitc_3),pch=1,col="blue")
```
In blue we have represented the predicted values according to the model logistic_3.
